#activate libraries
```{r}
library(RColorBrewer)
library(pheatmap)
library(MicrobiotaProcess)
library(phyloseq)
library(ggplot2)
library(ggtree)
library(ggtreeExtra)
library(tidytree)
library(ggstar)
library(forcats)
library(ggpubr)
library(coin)
library(ComplexUpset)
library(tidytree)
library(data.table)
library(DT)
library(tidyr)
library(ggVennDiagram)
library(aplot)
library(patchwork)
library(dplyr)
library(tibble)
library(ggvenn)
library(vegan)
```

#import data
```{r}
#Nanopore
meta_np <-read.csv("path/to/metadata_ont.csv", colClasses = "character", sep = ",")
meta_np <- meta_np[1:48, ]
meta_np$technology <- "ONT"
meta_np$sampleid <- paste(meta_np$sampleid, "ONT", sep = "_")
barcodes_np <- meta_np$barcode

emu_np <- read.csv("path/to/emu_ont.tsv", sep = "\t")
emu_np <- emu_np %>% filter(row.names(emu_np) %in% barcodes_np)
numeric_columns_np <- sapply(emu_np, is.numeric)
zero_sum_columns_np <- colSums(emu_np[, numeric_columns_np]) == 0
emu_np <- emu_np[, !zero_sum_columns_np]
col_names_np <- colnames(emu_np)
emu_np <- rownames_to_column(emu_np)

taxonomy_np <- read.csv("path/to/taxonomy_pac.tsv", sep = "\t")
taxonomy_np[taxonomy_np == ""] <- NA
taxonomy_np <- taxonomy_np %>% filter(species %in% col_names_np)

#PacBio
meta_pb <- read.csv("path/to/metadata_pac.csv", colClasses = "character", sep = ",")
meta_pb$technology <- "PacBio"
meta_pb$sampleid <- paste(meta_pb$sampleid, "PacBio", sep = "_")
barcodes_pb <- meta_pb$barcode

emu_pb <- read.csv("path/to/emu_10.tsv", sep = "\t")
emu_pb <- emu_pb %>% filter(row.names(emu_pb) %in% barcodes_pb)
#remove zero sum columns
numeric_columns_pb <- sapply(emu_pb, is.numeric)
zero_sum_columns_pb <- colSums(emu_pb[, numeric_columns_pb]) == 0
emu_pb <- emu_pb[, !zero_sum_columns_pb]
col_names_pb <- colnames(emu_pb) 
emu_pb <- rownames_to_column(emu_pb)

taxonomy_pb <- read.csv("path/to/taxonomy_10.tsv", sep = "\t")
taxonomy_pb[taxonomy_pb == ""] <- NA
#remove row with taxa with zero reads
taxonomy_pb <- taxonomy_pb %>% filter(species %in% col_names_pb)

emu_pb <- emu_pb[, c("rowname", taxonomy_pb$species)]
```

#merge and process data
```{r}
metadata <- rbind(meta_np, meta_pb)

taxonomy <- rbind(taxonomy_np, taxonomy_pb)
taxonomy <- unique(taxonomy)


colname_np <- colnames(emu_np)
colname_pb <- colnames(emu_pb)
common_taxa <- colname_np %in% colname_pb
match_emu <- colname_np[common_taxa]

emu <- full_join(emu_np, emu_pb, by = match_emu)
emu[is.na(emu)] <- 0
rownames(emu) <- emu$rowname
emu <- emu[, 2:length(colnames(emu))]

```

###procrusts test
```{r}
# Bray-Curtis + PCoA
dist_tax <- vegdist(emu, method = "bray")
pcoa_tax <- cmdscale(dist_tax, eig = TRUE, k = 2) 
X <- pcoa_tax$points  

# make dummy-matrix
Y <- model.matrix(~ technology - 1, data = metadata)  # remove intercept
colnames(Y) <- levels(factor(metadata$technology))    # clean names

test_result <- protest(X = X, Y = Y, permutations = 999)
print(test_result)
plot(test_result)
```



# import data as phyloseq object
```{r}
ps.taxonomy <- taxonomy
rownames(ps.taxonomy) <- ps.taxonomy$species

ps.meta <- metadata
rownames(ps.meta) <- ps.meta$barcode

ps <- phyloseq(otu_table(emu, taxa_are_rows = F), phyloseq::tax_table(as.matrix(ps.taxonomy[-1][c(6:1)])), sample_data(ps.meta[-1]))
```

# import data as MicrobiotaProcess
```{r}
ps %>% as.MPSE() -> mpse
mpse <- mpse %>% mp_filter_taxa(.abundance = Abundance, min.abun = 1, min.prop = 0.05)
mpse %<>% mp_rrarefy()
```

# alpha-diversity technology
```{r fig.height=4, fig.width=4}
mpse %<>% mp_cal_alpha(.abundance=RareAbundance)

alpha_plot_group <- mpse %>% 
    mp_plot_alpha(
        .group=technology, 
        .alpha=c(Shannon)
    )+
    scale_fill_brewer(palette = "Set1")+
    scale_color_brewer(palette = "Set1")+
  theme_bw() +
    theme(axis.text.x = element_text(angle = 45, size = 10, colour = "black", hjust = 1), 
          title = element_text(size = 13),
          strip.text = element_blank(),
          legend.position = "none",
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  labs(y = "Shannon index")
alpha_plot_group
```

#save alpha
```{r}
ggsave("path/to/alpha.pdf", plot = alpha_plot_group, device = "pdf", width = 4, height = 4)
```


# taxonomy barplots (genus, phylum)
```{r fig.height=6, fig.width=9}
mpse %<>%
    mp_cal_abundance( # for each samples
        .abundance = RareAbundance
    ) %>%
    mp_cal_abundance( # for each groups 
        .abundance=RareAbundance,
        .group=technology
    )


```
```{r fig.height=6, fig.width=6}
genus_barplot_1 <- mpse %>%
  mp_plot_abundance(
    width = 0.55,
    .abundance = RareAbundance,
    .group = technology,
    taxa.class = genus,
    rmun = TRUE,
    topn = 6,
    rm.zero = TRUE,
    plot.group = TRUE,
    relative = TRUE,
    force = TRUE
  ) 
plot_data <- ggplot_build(genus_barplot_1)
unique(plot_data[["data"]][[1]][["alluvium"]])

predefined_colors <- c(
  "Others" = "#54278F",
  "g__Brevitalea" = "#4D4D4D",
  "g__Solirubrobacter" = "#429375",
  "g__Baekduia" = "#883590",
  "g__Vicinamibacter" = "#F28629",
  "g__Bacillus" = "#E6DB71",
  "g__Gaiella" = "#1F78B4"
)

genus_barplot_1_1 <- genus_barplot_1 +
  scale_fill_manual(
    values = predefined_colors,
    labels = gsub("^g__", "", names(predefined_colors))
  ) +
  theme(
    legend.position = "none",
    legend.text = element_text(size = 16),
    legend.key.height = unit(0.4, 'cm'),
    legend.key.width = unit(0.3, 'cm'),
    axis.text.x = element_text(angle = 30, size = 20, colour = "black", hjust = 1),
    title = element_text(size = 24)
  ) +
  guides(fill = guide_legend(ncol = 1)) #+
  #xlab("Technology")


genus_barplot_1_1
```
#save bar
```{r}
ggsave("path/to/bar.pdf", plot = genus_barplot_1_1, device = "pdf", width = 6, height = 6)
```

#Venn new 
```{r fig.height=4, fig.width=6}
venn_plot <- ggvenn::ggvenn(list(ONT = taxonomy_np$species, PacBio = taxonomy_pb$species),text_size = 6, fill_color = c("red", "green"), fill_alpha = 0.6)
venn_plot
```
#save venn
```{r}
ggsave("path/to/venn.pdf", plot = venn_plot, device = "pdf", width = 6, height = 4)
```
#Venn genus new 
```{r fig.height=4, fig.width=6}
venn_genus <- ggvenn::ggvenn(list(ONT = unique(na.omit(taxonomy_np$genus)),
                                  PacBio = unique(na.omit(taxonomy_pb$genus))),
                             text_size = 6, 
                             fill_color = c("red", "green"),
                             fill_alpha = 0.6)
venn_genus
```
#save venn
```{r}
ggsave("path/to/venn_genus.pdf", plot = venn_genus, device = "pdf", width = 6, height = 4)
```

# beta-diversity PCA group
```{r}
mpse %<>% mp_decostand(.abundance=Abundance, method = "hellinger")
mpse %<>% mp_cal_pca(.abundance = hellinger)
```

# Permutational Multivariate Analysis of Variance (group)
```{r}
mpse %<>% mp_adonis(.abundance = hellinger, distmethod = "bray",
.formula = ~technology, permutation = 9999, action = "add")
mpse %>% mp_extract_internal_attr(name=adonis) %>% mp_fortify() 
```


# beta-diversity (Bray-Curtis dissimilarity distance) PCoA group
```{r fig.height=5, fig.width=5}
mpse %<>% mp_decostand(.abundance=Abundance)
mpse %<>% mp_cal_dist(.abundance=hellinger, distmethod="bray")
mpse %<>% mp_cal_pcoa(.abundance=hellinger, distmethod="bray")

bray_pcoa_group <- mpse %>%
    mp_plot_ord(
        .ord = pcoa, 
        .group = technology,
        .color = technology, 
        .size = 1.75,
        .alpha = 1,
        ellipse = TRUE,
        show.legend = F,
        show.side = F
    )+
    scale_fill_brewer(palette = "Set1", name = "Technology")+
    scale_color_brewer(palette = "Set1")+
    theme(axis.text.x = element_text(angle = 45, size = 10, colour = "black", hjust = 1), 
          title = element_text(size = 13),
          legend.position = "bottom") +
  annotate("text", x = 0.28, y = -0.32, label = "Pr(>F) = 1e-04", color = "red", size = 3) # add here the PERMANOVA p-value
bray_pcoa_group
```


# betta analysis save
```{r fig.height=12, fig.width=12}
ggsave("path/to/betta.pdf", plot = bray_pcoa_group, device = "pdf", width = 5, height = 5)
```
